---
title: "project"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dpi=120)
```

```{r load-libraries, message=FALSE}
library(dplyr)
library(ggplot2)
library(GGally)
library(inspectdf)
library(reshape2)
library(RColorBrewer)
library(caTools)
library(leaps)
library(glmnet)
library(pROC)
library(xtable)
library(stargazer)
library(corrplot)
```

## Exploratory data analysis

1. Load in the data
2. See what data looks like
3. Check shape of data

```{r echo=FALSE, results='hide',message=FALSE}
heart = read.csv('framingham.csv')
head(heart)
dim(heart)
summary(heart)
```

Checking number of NA rows

```{r echo=FALSE, results='hide',message=FALSE}
dim(heart[!complete.cases(heart),])

na.count = table(heart[!complete.cases(heart),]$TenYearCHD)

chd.count = table(heart$TenYearCHD)

# Proportion of NAs compared to original data
na.count/chd.count
```

Drop NA rows, check structure

```{r echo=FALSE, results='hide',message=FALSE}
heart = heart[complete.cases(heart),]
str(heart)
```

Copy data into new variable just for EDA

```{r echo=FALSE, results='hide',message=FALSE}
heart.eda = heart
heart.eda$TenYearCHD = factor(heart.eda$TenYearCHD, levels=c(0,1), labels=c("No", "Yes"))
heart.eda$male = factor(heart.eda$male)
heart.eda$education = factor(heart.eda$education)
heart.eda$currentSmoker = factor(heart.eda$currentSmoker)
heart.eda$BPMeds = factor(heart.eda$BPMeds)
heart.eda$prevalentStroke = factor(heart.eda$prevalentStroke)
heart.eda$prevalentHyp = factor(heart.eda$prevalentHyp)
heart.eda$diabetes = factor(heart.eda$diabetes)
```

Visualizing proportion of chd for each predictor

```{r echo=FALSE, results='hide',message=FALSE, fig.show = 'hide'}
#path = getwd()

#cairo_pdf("agevschd.pdf")
ggplot(heart.eda, aes(x=age, fill=TenYearCHD)) + 
  geom_density(alpha=0.4) + 
  labs(x="Age", fill="Developed heart disease")
#dev.off()


ggplot(heart.eda, aes(x=currentSmoker, fill=TenYearCHD)) + 
  geom_bar(alpha=0.4, position="fill") + 
  labs(x="Smokes", fill="Developed heart disease")

#cairo_pdf("totcholvschd.pdf")
ggplot(heart.eda, aes(x=totChol, fill=TenYearCHD)) + 
  geom_density(alpha=0.4) + 
  labs(x="Total Cholesterol", fill="Developed heart disease")
#dev.off()


ggplot(heart.eda, aes(x=male, fill=TenYearCHD)) + 
  geom_bar(alpha=0.4, position="fill") + 
  labs(x="Sex", fill="Developed heart disease")


#cairo_pdf("glucosevschd.pdf")
ggplot(heart.eda, aes(x=glucose, fill=TenYearCHD)) + 
  geom_density(alpha=0.4) + 
  labs(x="Glucose", fill="Developed heart disease")
#dev.off()

ggplot(heart.eda, aes(x=diaBP, fill=TenYearCHD)) + 
  geom_density(alpha=0.4) + 
  labs(x="Diastolic blood pressure", fill="Developed heart disease")


ggplot(heart.eda, aes(x=heartRate, fill=TenYearCHD)) + 
  geom_density(alpha=0.4) + 
  labs(x="Heart Rate", fill="Developed heart disease")


ggplot(heart.eda, aes(x=prevalentHyp, fill=TenYearCHD)) + 
  geom_bar(alpha=0.4, position="fill") + 
  labs(x="Prevalent Hypertension", fill="Developed heart disease")


ggplot(heart.eda, aes(x=diabetes, fill=TenYearCHD)) + 
  geom_bar(alpha=0.4, position="fill") + 
  labs(x="Diabetes", fill="Developed heart disease")

#cairo_pdf("sysbpvschd.pdf")
ggplot(heart.eda, aes(x=sysBP, fill=TenYearCHD)) + 
  geom_density(alpha=0.4) + 
  labs(x="Systolic Blood Pressure")
#dev.off()


```

Plotting each predictor against response variable

```{r echo=FALSE, results='hide',message=FALSE , fig.show = 'hide'}
for (feature in names(heart)){
  print(ggplot(heart, aes_string(feature, "TenYearCHD")) + geom_point())
}
  
```


```{r echo=FALSE, results='hide',message=FALSE}
create_heatmap = function(df) {
  
cor.mat = round(cor(df), 2)
cor.mat[upper.tri(cor.mat)] = NA
melted.cor = melt(cor.mat, na.rm = TRUE)
names(melted.cor) = c("X", "Y", "correlation")

ggplot(melted.cor, aes(X, Y, fill= correlation)) + 
  geom_tile() +
  scale_fill_distiller(limit = c(-1,1), palette = "RdBu") +
  theme_minimal() +
  geom_text(aes(X, Y, label = correlation), color = "black", size = 3) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank(),
        legend.justification = c(1, 0),
        legend.position = c(0.6, 0.7),
        legend.direction = "horizontal")+
        guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                        title.position = "top", title.hjust = 0.5))
}
```

```{r echo=FALSE, results='hide',message=FALSE, fig.show = 'hide'}
create_heatmap(heart)

cairo_pdf("correlation.pdf", width=8)

corrplot(cor(heart),
         method="color",
         tl.col="black",
         tl.cex=1.2,
         tl.srt=55,
         number.cex=.5,
         cl.cex=1.2,
         cl.pos = "b",
         mar=c(0,0,1,0))
dev.off()
```

Start removing variables

```{r}
heart2 = heart

for(i in 1:length(colnames(heart2))) {
  colnames(heart2)[i] = tolower(colnames(heart2)[i]) 
}

colnames(heart2)[1] = "ismale"

heart2 = subset(heart2, select= -c(diabp))

# Only 21 people had stroke
table(heart2$prevalentstroke)

heart2 = subset(heart2, select= -c(prevalentstroke, currentsmoker, diabetes, prevalenthyp))

create_heatmap(heart2)
```


```{r results='hide',message=FALSE}
heart2$education = factor(heart2$education)

str(heart2)
```

```{r results='hide',message=FALSE}
heart3 = subset(heart2)

set.seed(11)
split.mask = sample.split(heart3$tenyearchd, SplitRatio = .8)
train = subset(heart3, split.mask == TRUE)
test  = subset(heart3, split.mask == FALSE)

```

## Modeling

```{r results='hide',message=FALSE}
model1 = glm(tenyearchd ~ 1, data=train, family = "binomial")

step.forward = step(model1, 
                    scope=list(lower= ~ 1, 
                               upper = ~ ismale + age + education +
                                 cigsperday + bpmeds +
                                 totchol + sysbp + bmi + heartrate +
                                 glucose), direction="forward")
```

```{r}
summary(step.forward)
anova(step.forward, test="Chisq")
```

```{r}
model2 = glm(tenyearchd ~ age + sysbp  + ismale + 
               cigsperday + glucose, 
             data=train, 
             family=binomial)

summary(model2)

anova(model2, test="Chisq")
```

prevalenthyp and bpmeds don't seem to be significant predictors at all according to the analysis of deviance table. Adding them to any reduced model does not make them significant.

```{r results='hide',message=FALSE}
model3 = glm(tenyearchd ~ ismale + age + education +
                                 cigsperday + sysbp + bmi + heartrate +
                                 glucose + bpmeds, data=train, family = "binomial")

step.backward = step(model3, 
                    scope=list(lower= ~ 1, 
                               upper = ~ .), direction="backward")

```

```{r}
summary(step.backward)
anova(step.backward, test="Chisq")
```

Calculate potentially useful statistics and print them. Also output a ROC plot.
```{r}
calculate_acc = function(model, threshold) {
    model.probs = predict(model, test, type="response")
    model.pred = ifelse(model.probs >= threshold, 1, 0)
    model.actual = test$tenyearchd
    
    conf.mat = table(model.pred, model.actual)
    conf.mat
    
    # Proportion of positive predictions that were correct
    precision = conf.mat[4]/(conf.mat[4] + conf.mat[2]) 
    precision
    
    # Proportion of actual positives identified correctly
    recall = conf.mat[4]/(conf.mat[4] + conf.mat[3])
    recall
    
    # Proportion predicted correctly
    accuracy = (conf.mat[4] + conf.mat[1])/(conf.mat[1]+conf.mat[2]+conf.mat[3]+conf.mat[4])
    paste(accuracy)
    
    # Accuracy if just guessed 0 for all cases
    baseline = (conf.mat[1] + conf.mat[2])/(conf.mat[1]+conf.mat[2]+conf.mat[3]+conf.mat[4])
    par(pty="s")
    plot(roc(model.actual,model.probs), grid=TRUE, print.auc=TRUE,
         cex.lab=2.,
         cex.axis=2.,
         cex.sub=2.,
         print.auc.cex=2.)
    
    print(conf.mat)
    print(accuracy)
    print(baseline)
    
}

calculate_acc(model2, 0.35)
```

Add some points to the roc plot

```{r}
par(pty="s")
calculate_acc(model2, 0.35)
points(.992, .144, pch=19)
text(.992, .144, labels="(0.5)", pos=4, cex=2)
points(.945, .356, pch=19)
text(.945, .356, labels="(0.35)", pos=4, cex=2)
```

Make nice tables for the models (outputs latex code)

```{r, results='asis'}
stargazer(step.forward, step.backward, model2,
          title="Continuous Models",
          column.labels = c("Forward Stepwise", "Backward Stepwise", "Significant Subset"),
          model.numbers = FALSE,
          header = FALSE,
          single.row = TRUE)
```


```{r, results='asis'} 
#xtable(summary(model2), display=rep('g',5))
xtable(anova(step.forward, test="Chisq"), display=c("d", "d","f", "d", "f", "g"))
```
